Multiscale Modeling of the Paradoxical Roles of the Immune System in EMT-Mediated Cancer

### 2019-03-05 Paper/Paper.tex
add refs

\documentclass[11pt, a4paper, preprint]{article}
\usepackage{outline}
\usepackage{float}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{cite}
\graphicspath{ {Figs/} }

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\newcommand{\tcr} { \textcolor{red} }

\title{Multiscale Modeling of the Paradoxical Roles of the Immune System in EMT-Mediated Cancer}
\author{}

@@

\section{Introduction}
Immunotherapy is slowly revolutionizing cancer therapy. New breakthroughs in the field begin to produce new therapies that have had a significant impact on patient health and survival\cite{pardoll2012blockade}\cite{restifo2012adoptive}.
% A More examples? Reviews or non-reviews? Explain some of the examples?
%AM I will add a few 
%There are myriad untapped avenues by which scientists can further explore the role of the immune system in regulating and combatting cancer;
In addition to direct cell-cell interactions between immune and cancer cells, the means by which immune cells interact with and modulate the tumor microenvironment (TME) is an important \tcr{research} area.
Immune cells can be understood as one constituent of the TME: responding to the environment's inflammatory signals, fighting the mutated cells present, and then restoring homeostasis in that local context\cite{de2006paradoxical}.
Understanding each step in this immune cascade in the context of the TME will be necessary in order to fully explore and exploit all options of immunotherapy.

\tcr{Still need to address Qing's comments on this paragraph:}

The tumor interacts with the immune system by initiating the sequence of events that results in the activation of the adaptive immune response.
It accomplishes this by presenting antigens that innate immune cells recognize and carry to lymph nodes to activate, among other things, T cells.
Also, tumor cells engage in certain cellular processes that shape the TME and thus impact the immune system.
In particular, tumor cells release transforming growth factor beta (TGF-$\beta$) that creates a tumor-supportive environment by shifting the immune balance to being immunosuppressive via enhancing activation of Treg cells.

@@

Epithelial-to-mesenchymal transition (EMT) describes the (reversible) cell plasticity process by which epithelial cells transition into mesenchymal cells.
While the epithelial cells often stay bound together in layers, mesenchymal cells have gained motility and even stem-like properties\cite{nieto2016emt}.
Recent work has revealed the intriguing possibility that there are intermediate cell states along the EMT spectrum with each state having its own unique properties\cite{hong2015ovol2}.
Such intermediate states have the potential of being connected to tumor initiation, progression, and metastasis\cite{nieto2016emt}.
While there might be many other characteristics distinctive of mesenchymal cells that need further exploration, two qualities of mesenchymal cells are particularly relevant in the context of cancer and the immune system:
1) mesenchymal cells are less susceptible to immune clearance\cite{terry2017new} and
%
2) mesenchymal cells proliferate at a slower rate. \tcr{Ref}
 %A Adam said he had a reference in mind for this.

@@

\bibliography{mybib}{}
\bibliographystyle{siam}

>>>>

\documentclass[11pt, preprint]{article}
\usepackage{outline}
\usepackage{float}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{cite}
\graphicspath{ {Figs/} }

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}

%\usepackage[ ]{natbib} 
\newcommand{\tcr} { \textcolor{red} }

% A multiscale model of EMT-mediated cancer reveals 
\title{Multiscale Modeling of the Paradoxical Roles of the Immune System in EMT-Mediated Cancer}
\author{Daniel R. Bergman$^{1}$,
Matthew Karikomi\,$^{1}$,
Qing Nie\,$^{1,2,*}$
and Adam L. MacLean\,$^{3,*}$
}
\affil{
  $^1$Department of Mathematics, University of California, Irvine, and \\
  $^2$Department of Cell and Developmental Biology, University of California, Irvine, Irvine, CA 92697, USA \\
  $^3$Department of Biological Sciences, University of Southern California, Los Angeles, CA 90089, USA \\
  $^*$Correspondence: macleana@usc.edu (ALM); qnie@uci.edu (QN).
}

@@

\section{Introduction}
The immune system affects tumorigenesis and cancer progression in myriad intricate ways. New breakthroughs in the field of immunotherapy are beginning to produce new therapies that have had a significant impact on patient health and survival \cite{pardoll2012blockade,restifo2012adoptive}. Beyond direct tumor-immune cell interactions, the immune system also modulates the tumor microenvironment (TME), and indeed, immune cells can be considered as a part of the TME: responding to local inflammatory signals, targeting mutated cells, and potentially eradicating tumor cells and restoring homeostasis in a local context \cite{de2006paradoxical}.
\par 
Tumor cells also interact with the immune system by initiating a sequence of events that results in the activation of the adaptive immune response. This is accomplished by the presentation of antigens that innate immune cells recognize and carry to lymph nodes to activate T cells (and other components) \cite{schreiber11_cancer}. The tumor also engages in cellular processes that indirectly shape the TME, for example by releasing transforming growth factor beta (TGF-$\beta$), thus shifting the TME towards a tumor-supportive environment by enhancing the immunosuppression via Treg cell activation \cite{schreiber11_cancer}.
\par
Epithelial-to-mesenchymal transition (EMT) describes reversible process by which cells with an epithelial phenotype transition into cells with a mesenchymal phenotype.
While epithelial cells are in-part defined by their adhesion to neighbors, mesenchymal cells show much greater ranges of motility, and perhaps stem-like properties \cite{nieto2016emt}, although controversy regarding `stemness' and EMT remains \cite{nie18_stem, sha19_intermediate}.  
Recent work has shown that -- rather than being a binary process -- at least two stable intermediate states exist on the EMT axis \cite{hong2015ovol2, jolly15_coupling}. Ongoing investigations into the plasticity and stability of EMT overlap with studies of discrete vs. continuous processes elsewhere, such as in cell differentiation \cite{moris16_transition}. Moreover, intermediate states have emerged as a central mechanisms by which cell fates (and the noise inherent within in them) can be controlled \cite{maclean18_exploring, ta16_controlling, rackauckas18_meanindependent}

While EMT has for some time been assumed to impact metastasis, more recently EMT-related phenotypes have been linked to other aspects of both tumor initiation and progression \cite{nieto2016emt}. In addition, TGF-$\beta$ is a master regulator of EMT, thus linking it closely with the signaling networks involved in immune responses [REF].

@@

While there might be many other characteristics distinctive of mesenchymal cells that need further exploration, two qualities of mesenchymal cells are particularly relevant in the context of cancer and the immune system:
1) mesenchymal cells are less susceptible to immune clearance\cite{terry2017new} and
%
2) mesenchymal cells proliferate at a slower rate. \tcr{Ref}

% model cancer-immune 
%\cite{West19_JCOClinicalCancerInformatics}

@@

\bibliography{mybib,amrefs}
\bibliographystyle{plos2015}
